---
title: "Getting start to feasibilityR"
output: rmarkdown::html_vignette
author: Saavedra et al. 2023
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup,echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

**Introduction**

The `feasibilityR` package provides a general interface to estimate the feasibility of a given community given different types of ecological communities and structure of species interactions. The feasibility domain mathematically defined as:

INCUDE FORMULA HERE

defines the opportunities for species to coexist. A larger feasibilty domain implies greater opportunities for species to coexist because it the community can accomodate a larger differences in intrinsic growth rate between species compatible with their coexistence. 

To calculate the feasibility domain of a given community with S species 

```{r}
library(feasibilityR)
#ADD FUNCTIONS OR GRAPHS HERE, COPY IT AS MANY TIMES AS NEEDED
```

**References**

ADD references here. 


