---
title: "Getting start to feasibilityR"
output: rmarkdown::html_vignette
author: Saavedra et al. 2023
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup,echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

**Introduction**

The `feasibilityR` package provides a general interface to estimate the feasibility of a given community given different types of ecological communities and structure of species interactions. 

Departing from a general Lotka-Volterra (LV) model that describes reasonable well the population dynamics of interacting species, 

$$ \frac{dN_{i}}{dt} =N_{i}\left(r_{i} + \sum_{j=1}^{S} \alpha_{ij}N_{j}\right)$$

the feasibility domain of a given community is mathematically defined as:

INCUDE FORMULA HERE

\Omega$ defines the opportunities for species to coexist. A larger feasibility domain implies greater opportunities for species to coexist because the community can accommodate larger differences in intrinsic growth rate between species compatible with their coexistence. 

To calculate the feasibility domain of a given community with S species, we need a matrix that contains the strength and sign of species interactions. Elements in the diagonal will represent intraspecific interactions while interspecific interactions are contained in the off-diagonal elements. Such interaction describes the per-capita effect of species i on j (alphaij) or 

Several matrices can be included obtained from experimental or field data. 




```{r}
library(feasibilityR)
#ADD FUNCTIONS OR GRAPHS HERE, COPY IT AS MANY TIMES AS NEEDED
```

**References**

ADD references here. 


