---
title: "Getting start to feasibilityR"
output: rmarkdown::html_vignette
author: Saavedra et al. 2023
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup,echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

**Introduction**

The `feasibilityR` package provides a general interface to perform a feasibility analysis of ecological systems following general Lotka-Volterra (gLV) dynamics.

Feasibility ($$\Omega$$): the size of the feasibility domain compatible with the possibility of long-term persistence of a given set of species $$\mathcal{C}$$ from a given community (pool) of species $$\mathcal{S}$$. The size is estimated by the normalized surface area of a unit sphere (i.e., also known as the normalized solid angle). Note that feasibility (size of the feasibility domain) is not directly a probability measure. To do so, one needs to normalize the feasibility such that values lie within $[0,1]$ according to the characteristics of the community.

Formally, the feasibility domain of any set $$\mathcal{C} \subseteq \mathcal{S}$$ under gLV dynamics can be expressed as

$$D(\mathcal{C}) = \left\{ \sum_{j \in \mathcal{C}} N_j^* \mathbf{a}_j - \sum_{k \in \mathcal{S}/\mathcal{C}} N_k^* \mathbf{e}_k : \mathbf{A} =\left[\mathbf{a}_1\ \mathbf{a}_2\ \cdots \  \mathbf{a}_{S}\right], N_j^*,  N_k^* \in \mathbb{R}_{>0} \right\} \in \mathbb{R}^{S}$$.


```{r}
library(feasibilityR)

matA <- generate_inte_rand(4, 1, 1, 0, "norm")  ## Generate a random matrix
matB <- generate_inte_gs(4, 1, 1, 0, "norm") ## Generate a globally stable random matrix
r <- matA %*% c(runif(4, 0, 1)) ## This is the direction of the r-vector (assuming it is known). This is only needed to calculate resistance and recovery

feasibility_community(matA) ## If the user is interested in the species-specif measure, then the user needs to calculate raise this outcome to the power 1/|S|, where |S| is the matrix dimension
feasibility_community(matB)

feasibility_overlap(matA, matB) ## Matrices need to have the same dimension

feasibility_asymmetry(matA) ## This is a standard deviation. The larger the outcome, the larger the assymetry

feasibility_partition(matB) ## The feasibility of all the 2^|S| possible combinations

feasibility_species(matB, sp = 2) ## The sum of feasibility regions whith species i

feasibility_resistance_full(matA, r) ## This is a distance from a a given r-vector to all possible |S-1| regions. The larger the outcome, the larger the resistance

feasibility_resistance_partial(matA, r) ## This is a distance from a a given r-vector to all possible vertices. The larger the outcome, the larger the resistance

feasibility_recovery(matA, r, type = "full") ## This is the largest eigenvalue (notice is negative)

feasibility_recovery(matA, r, type = "part") ## This is the second smallest eigenvalue (notice is negative)

feasibility_contribution(matB, sp = 2) ## This is a ratio between the feasibility with |S| species and with |S-1| species. Otcomes above (resp. below) 1 mean a positive (resp. negative) contribution

feasibility_invasion(matB, invader = 1) ## This is the feasibility region for invasion

```

**References**

[1] Chuliang Song, Rudolf P Rohr, and Serguei Saavedra. “A guideline to study the feasibility
domain of multi-trophic and changing ecological communities”. In: Journal of Theoretical
Biology (2018).

[2] Serguei Saavedra, Lucas P Medeiros, and Mohammad AlAdwani. “Structural forecasting of
species persistence under changing environments”. In: Ecology letters (2020).

[3] Andrea Tabi et al. “Species multidimensional effects explain idiosyncratic responses of com-
munities to environmental change”. In: Nature Ecology & Evolution (2020).

[4] Ignasi Bartomeus et al. “Experimental evidence of the importance of multitrophic structure
for species persistence”. In: Proc. Nat. Acad. of Sci. USA 118.12 (2021).

[5] Lucas P Medeiros, Chuliang Song, and Serguei Saavedra. “Merging dynamical and structural
indicators to measure resilience in multispecies systems”. In: Journal of Animal Ecology (2021).

[6] Jie Deng, Washington Taylor, and Serguei Saavedra. “Understanding the impact of third-
party species on pairwise coexistence”. In: PLOS Computational Biology (2022).

[7] Jie Deng et al. “On the limits to invasion prediction using coexistence outcomes”. In: bioRxiv
(2023). doi: 10.1101/2023.03.23.533987.
